Quantum Interval-Valued Probability: Contextuality and the Born Rule

### 2017-08-15 prop_letter.tex
prop_letter.tex Add citations...

In the long version of this paper, we establish that QIVPMs become
classical IVPMs when the space of events includes only mutually
commuting events. The idea of the proof is as follows. Given a
subspace of commuting events $\events_C$, we can construct a partial
orthonormal basis by diagonalization and complete it to a full
orthonormal basis $\overline{\events_C}$. We then build a bijection
between the QIVPM on the set of projectors associated with this basis
and the set of classical events corresponding to this basis. This
correspondence leads to a crucial property in our setting.

\begin{thm}\label{thm:Shapley}
  For a QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, if a subspace
  of events~$\events_C\subseteq\events$ commutes, then
  $\coreBorn\left(\bar{\mu},\events_C\right)\ne\emptyset$.
\end{thm}

@@

\begin{example}\label{ex:three-dimensional-three-value}
  Given a three-dimensional Hilbert space $\Hilb$ with an orthonormal
  basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $, let
  $\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
  $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
  $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\,,
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\,,
\end{aligned}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}
It is straightforward to verify that $\bar{\mu}$
is a QIVPM. Now let $S_{1}=\left\{ \proj{0},\proj{\ps}\right\} $.
Because $\bar{\mu}(\proj{0})=\bar{\mu}(\proj{\ps})=\imposs$, we have
$\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=0$
for any state~$\rho\in\coreBorn\left(\bar{\mu},S_{1}\right)$.
Hence, $\rho$ cannot be any state other than $\ket{2}$, the only
pure state orthogonal to both $\ket{0}$ and $\ket{\ps}$, and it
is easy to verify $\proj{2}$ is actually in $\coreBorn\left(\bar{\mu},S_{1}\right)$.
Similarly, let $S_{2}=\left\{ \proj{0},\proj{\ps'}\right\} $. We
have $\coreBorn\left(\bar{\mu},S_{2}\right)=\left\{ \proj{1}\right\} $.
Therefore, 
\begin{equation}
\coreBorn\left(\bar{\mu},\events\right)\subseteq\coreBorn\left(\bar{\mu},S_{1}\right)\cap\coreBorn\left(\bar{\mu},S_{2}\right)=\emptyset
\end{equation}
Thus, there is no state that is consistent with the QIVPM on all projectors~$\events$,
so Gleason's theorem does not hold for at least this example.
\end{example}

>>>>

In the long version of this paper~\cite{HOSTunpublished}, we establish that
QIVPMs become classical IVPMs when the space of events includes only mutually
commuting events. The idea of the proof is as follows. Given a
subspace of commuting events $\events_C$, we can construct a partial
orthonormal basis by diagonalization and complete it to a full
orthonormal basis $\overline{\events_C}$. We then build a bijection
between the QIVPM on the set of projectors associated with this basis
and the set of classical events corresponding to this basis. This
correspondence leads to a crucial property in our setting.

\begin{thm}\label{thm:Shapley}
  \cite{Shapley1971,GilboaSchmeidler1994,Grabisch2016}~For a
  QIVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}$, if a subspace of events~$\events_C\subseteq\events$ commutes, then
  $\coreBorn\left(\bar{\mu},\events_C\right)\ne\emptyset$.
\end{thm}

@@

\begin{example}\label{ex:three-dimensional-three-value}
  Given a three-dimensional Hilbert space $\Hilb$ with an orthonormal
  basis $\left\{ \ket{0},\ket{1},\ket{2}\right\} $, let
  $\mathscr{I}_{1}=\left\{ \necess,\imposs,\unknown\right\} $,
  $\ket{\ps}=\left(\ket{0}+\ket{1}\right)/\sqrt{2}$,
  $\ket{\ps'}=\left(\ket{0}+\ket{2}\right)/\sqrt{2}$, and
\begin{subequations}
\begin{eqnarray}
 &  & \begin{aligned}\imposs & =\bar{\mu}(\mathbb{0})=\bar{\mu}(\proj{0})\\
 & =\bar{\mu}(\proj{\ps})=\bar{\mu}(\proj{\ps'})\,,
\end{aligned}
\\
 &  & \begin{aligned}\necess & =\bar{\mu}(\mathbb{1})=\bar{\mu}(\mathbb{1}-\proj{0})\\
 & =\bar{\mu}(\mathbb{1}-\proj{\ps})=\bar{\mu}(\mathbb{1}-\proj{\ps'})\,,
\end{aligned}
\\
 &  & \unknown=\bar{\mu}(P)\qquad\textrm{otherwise.}
\end{eqnarray}
\end{subequations}
It is straightforward to verify that $\bar{\mu}$ is a QIVPM. Then we will
prove $\coreBorn\left(\bar{\mu},\events\right)$ is an empty set by
contradiction. Suppose there is a
state\nb$\rho=\sum_{j=1}^{N}q_{j}\proj{\phi_{j}}\in\coreBorn\left(\bar{\mu},\events\right)$,
where $\sum_{j=1}^{N}q_{j}=1$ and $q_{j}>0$. Since $\mu_{\rho}^{\mathrm{B}}(P)\in\bar{\mu}(P)$
for any\nb$P$, the state\nb$\rho$ must satisfy
\begin{equation}
\mu_{\rho}^{\mathrm{B}}(\proj{0})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps})=\mu_{\rho}^{\mathrm{B}}(\proj{\ps'})=0
\end{equation}
which implies $\ip{\phi_{j}}{0}=\ip{\phi_{j}}{\ps}=\ip{\phi_{j}}{\ps'}=0$
for all $j$. However, the last equation never holds. 
Thus, there is no state consistent with the QIVPM on all 
projectors\nb$P\in\events$, so Gleason's theorem does not hold for at least this
example.
\end{example}
